---
title: "PDA_NYC_BIKES"
output: html_document
date: '2022-04-19'
---

```{r setup, include=FALSE} 
knitr::opts_chunk$set(warning = FALSE, message = FALSE) 
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
library(tsibble)
library(tsibbledata)
library(tidyverse)
library(janitor)
library(lubridate)
library(ggplot2)
library(leaflet)
library(sf)
library(rgeos)
library(rnaturalearth)
library(rnaturalearthdata)
```


```{r}
nyc_bikes_df <- clean_names(nyc_bikes)
```

1.1
citi bikes are a company that operate in new york city who provide a service of publicly accessible bicycles from stations which can then be returned to a simmilar station at a different location makeing it an ideal mode of transport for fast, eco freindly, inexpensive one way travel.

1.2
this analysis will analyse the target demographic for citi bikes to advartise and cater to to maximise their frequent customer satisfaction.
aswell as identify peak times of usage for the bikes wherein more resources will be required for an efficant service.

1.7
the data is processed before visualisation to make it easier to understand, the days of the week are ordered chronologicaly and impossible ages are removed to prevent the integrity of the data being compromised. the age of the person is calculated from the birth year and the year of the reccord.

1.4
the average age for a user is between 25 and 35, while an anomalous stand out group does seem to exist, the gender imporbability seen here calls into the question the reliability of this data for these factors.

the ratios of genders using the service appear to be consistent with males being greater users of the service by around double.
this can affect the marketing campaigns and market audience decisions going forward.
a massive portion of the unknown genders is unknown, however this is magnified in a particular age group which is a statistical improbability.
its very possible the data on these people is inaccurate.

1.6
it is also possible that a system default has been set for unidentified users. this should be removed if it is the case as unlabled data is more usefull when classified as unknown rather than skewing the remainder of the valid data.

```{r}
nyc_bikes_df <- nyc_bikes_df %>%
  mutate(day = weekdays(start_time), month = months(start_time), year = year(start_time)) %>%
  mutate(age = year - birth_year) %>%
  mutate(age = ifelse(age>100, NA, age))


nyc_bikes_df$day <- ordered(nyc_bikes_df$day, levels=c("Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday", "Sunday"))
nyc_bikes_df$month <- ordered(nyc_bikes_df$month, levels=c("January", "February", "March", "April", "May", "June", "July", "August", "September", "October", "November", "December"))

```


all bike usage over time

```{r}
ggplot(nyc_bikes_df, aes(x=start_time, y=..density..)) +
  geom_histogram(bins = 52, colour = "black") + 
  geom_density(colour = "red") +
  xlab("date")
```

this yearly diagram of service use demonstrates a seasonal pattern, however more data will be required to confirm this theory.
current marketing and business decisions should be conducted with expecte high use in july and august and limmited use over the winter




peak bike usage times

```{r}
ggplot(nyc_bikes_df, aes(x=month, fill = day)) +
  geom_bar() + 
  xlab("month")
```


the highest usage for the service is dureing the months of july and august likely due to the good weather.
the most popular days for usage however are weekdays which possibly suggests there is a market for commuters using this service for transport to work



bike usage by age and gender

```{r}
ggplot(nyc_bikes_df, aes(x=age, fill = gender)) +
  geom_bar() + 
  xlab("age") +
  theme_bw()
```






```{r}
bike_station_data <- select(nyc_bikes_df, start_station, start_long, start_lat) %>%
distinct(start_station, .keep_all = TRUE)

leaflet(bike_station_data) %>%
  addTiles() %>%
  setView(-74.06, 40.72, zoom = 13) %>%
  addCircleMarkers(lng = ~start_long, lat = ~start_lat, 
             popup = ~start_station)
```

this plot shows the geographical position between the stations.
there are 52 active stations in the area covered by the data.
the density of stations near jersey city is greater and so the possibility for inefficiancy increases

the company should focus advertiseing on a demographic of 35 year old men and prepare for high usage in the months around august.


future work on this dataset should include an analysis on which stations are most used and the distances of each trip using the station locations coordinates.
      this can also be broken down into the average distance covered by individual bikes. this information can be used to rotate bikes between high use and low use stations.
      or position the bikes at the stations such that the bikes of low use become more available to distruibute the bikes usage.

