---
title: "R Notebook"
output: html_notebook
---

```{r}
library(palmerpenguins)
library(tidyverse)
library(ggplot2)
library(janitor)
library(e1071)
library(gginference)

penguins <- palmerpenguins::penguins 
penguins <-  clean_names(penguins)

summary(penguins)
```

```{r}
ggplot(penguins, aes(x = species, y = bill_length_mm)) + 
    geom_boxplot()
```

the Adelie penguin has the shortest beak length by far, the chinstrap has the longestet beak lenth but there is significant overlap with the gentoo pengiun.


```{r}
penguins %>%
  filter(!is.na(bill_length_mm)) %>%
  group_by(species) %>%
  summarise(mean(bill_length_mm), median(bill_length_mm), sd(bill_length_mm), skewness(bill_length_mm))
```

the gentoo penguin has data moderatly skewwed to the right, the other two penguines have more ballenced data. 
the standered deviations show that the data between penguins is simmilarly distributed.

```{r}
chinstrap <-  penguins %>%
   filter(!is.na(bill_length_mm), species == "Chinstrap")
gentoo <-  penguins %>%
   filter(!is.na(bill_length_mm), species == "Gentoo")
  
px = t.test(chinstrap$bill_length_mm, gentoo$bill_length_mm)
px

ggttest(px)
```

P value is less than 0.05 and thus we reject the null hypothesis.
the T value is greater than 2 so we can infer there is a significant difference betweeen the data.



