---
title: "Analysis_Meteorites"
output: html_notebook
---
3.1 laoding data
loading in clean datafile and reqired librarys

```{r}
library(tidyverse)
raw_clean_data = read.csv("meteorite_landings_CLEAN.csv")
```


3.4 & 3.5 below
wranglaing data & statistical analysis

10 largest metiorites within this data
```{r}
head(arrange(raw_clean_data, desc(mass_g)), 10)
```

avarage mass of falling meteorites vs found meteorites
```{r}
falling_meteorites = mean(filter(raw_clean_data, fall == "Fell")$mass_g)
found_meteorits = mean(filter(raw_clean_data, fall == "Found")$mass_g)
print(paste("avarage mass (g) of fallen meteorites: ", falling_meteorites))
print(paste("avarage mass (g) of found meteorites: ", found_meteorits))
```

number of meteorites annually since the year 2000
```{r}
after_2000 = filter(raw_clean_data, year >= 2000)
table(after_2000$year)
```

function get_meteorite_count version 1, takes in data and year 

```{r}
get_meteorite_count = function(input_data, input_year){
  filtered_data = filter(input_data, input_year == year)
  print(paste("in", input_year))
  if(nrow(filtered_data > 0)){
    print(paste("there were", nrow(filtered_data), "meteorites"))
  } else {
    print("As far as we know, there were no meteorites that year")
  }
}
```

4.1 test plans

test plans
current function get_meteorite_count can be tested by entering avaialbe test data with known values and observing the output.
the function also suffers from several weaknesses such as no ability to identify an invalid input
the function also does not output a value for use in other functions and only prints to the console, adding return data functionality would allow this function to be used in tandom for data analysis
assertive programming required.
for testing, 
insert invalid data, 
pass a string into the year parameter.
ensure a date before records began 

4.6 performing test

```{r}
library(testthat)

test_that("function (get_meteorite_count) gives expected outputs", {
  expect_error(get_meteorite_count(raw_clean_data, "2000"))
  expect_error(get_meteorite_count(raw_clean_data, "asdf"))
  expect_error(get_meteorite_count(raw_clean_data, TRUE))
  expect_error(get_meteorite_count(c(1,2,3,4), 2000))
  })
```

tests failed, charichters, strings and logicals in the year parametr should throw errors
4.6 revising code

```{r}
refactor_get_meteorite_count = function(input_data, input_year){
  if(!is.numeric(input_year)){
    stop("the year is not a valid numeric value")
  }
  if(!is.data.frame(input_data)){
    stop("the input is not a data frame")
  }
  filtered_data = filter(input_data, input_year == year)
  print(paste("in", input_year))
  if(nrow(filtered_data > 0)){
    print(paste("there were", nrow(filtered_data), "meteorites"))
  } else {
    print("As far as we know, there were no meteorites that year")
  }
  return(nrow(filtered_data))
}
```

new function testing

```{r}
library(testthat)

test_that("function (get_meteorite_count) gives expected outputs", {
  expect_error(refactor_get_meteorite_count(raw_clean_data, "2000"))
  expect_error(refactor_get_meteorite_count(raw_clean_data, "asdf"))
  expect_error(refactor_get_meteorite_count(raw_clean_data, TRUE))
  expect_error(refactor_get_meteorite_count(c(1,2,3,4), 2000))
  })
```

